---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
require(dplyr)
require(data.table)
require(ggplot2)
require(lattice)
require(latticeExtra)
require(knitr)
```

## Loading and preprocessing the data
```{r, echo=TRUE}
##Load & process the data
activ_data <- read.csv("activity.csv")
activ_data$date <- as.Date(activ_data$date, "%Y-%m-%d")
```



##Calculate the total number of steps taken per day
```{r, echo=TRUE}
steps_day <- aggregate(steps ~ date, activ_data, sum)
steps_day
```



##Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
hist(steps_day$steps)
```

##Summary statistics to claculate the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
summary(steps_day$steps)

```

## 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r, echo=TRUE}
dt = data.table(activ_data) ###Copy original data and convert to data table
dt[, steps.mean := mean(steps, na.rm = T), by = interval] ###Add columen with mean steps acrose all days by time interval
interval_ave_max <- filter(dt, max(steps.mean)  == steps.mean) ### Use dplyr to subset table by max mean steps
unique(interval_ave_max$interval) ### 5-minute interval corasponding to max mean
```

## Calculate and report the total number of missing values in the datase
```{r, echo=TRUE}
sum(is.na(activ_data))
```

## Imputing missing values
## New dataset with missing values filled in with mean for that 5-minute interval
```{r, echo=TRUE}
no_missing_values  <- as_data_frame(dt) ###Transform table in to dataframe

for(i in 1:17568) { ### for loop to replace missing values
  if(is.na(no_missing_values$steps[i])){
    no_missing_values$steps[i] = no_missing_values$steps.mean[i]
  }else(no_missing_values$steps[i] = no_missing_values$steps[i])
}
```




```{r, echo=TRUE}
##Calculate the total number of steps taken per day
steps_day_2 <- aggregate(steps ~ date, no_missing_values, sum)  

##Make a histogram of the total number of steps taken each day
hist(steps_day_2$steps)

##Compare with summary statistics after imputing missing values
summary(steps_day_2$steps)
summary(steps_day$steps)


```

##  Differences in activity patterns between weekdays and weekends
```{r, echo=TRUE}
##Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
day_of_week <- weekdays(no_missing_values$date)
day_of_week <- as.data.frame(day_of_week) ### convert to data dataframe 
week_day_end <- gsub("Saturday|Sunday", "weekend", day_of_week[,1]) ### serch for days corisponding to weekend and replace 
week_day_end <- gsub("Monday|Tuesday|Wednesday|Thursday|Friday", "weekday", week_day_end) ### serch for days corisponding to weekday and replace 

with_Day_column <-  data.frame(no_missing_values, week_day_end) ### bind sepret data frames in to one table
#with_Day_column <- as.data.frame(with_Day_column)

with_Day_column_2 <- aggregate(steps ~ week_day_end+interval, with_Day_column, mean) ### average number of steps across all weekday days or weekend days

ggplot(with_Day_column_2, aes(x=interval , y=steps, color = week_day_end))+
  geom_line(size=1)+
  facet_grid( week_day_end~. )+
  ggtitle("Average Number of Steps in Five Minute Intervals")

xyplot(steps ~ interval | week_day_end,  data = with_Day_column_2,
       ylab = "Mean Steps", xlab = "Interval", main = "Average Number of Steps in Five Minute Intervals",
       type = "l", lty = 1, col = "black", layout = c(1, 2), lattice.options = theEconomist.opts())

```

